\clearpage
\newgeometry{hmargin=3.5cm,vmargin=5.3cm,head=3.5cm,foot=3.2cm}
\begin{centering}
\textbf{ACKNOWLEDGEMENTS}\\
\vspace{\baselineskip}
\end{centering}

As I complete my master’s degree, I am grateful to have been surrounded by wonderful people, which allowed me to enjoy a happy graduate school life. 
I would like to express my heartfelt thanks to the precious individuals who helped me finish this journey well.

First, I extend my deep gratitude to my advisor, Professor Seong Lyun Kim, for his invaluable guidance and support throughout my research. 
His expertise and encouragement have been instrumental in shaping my academic journey. 
I am also thankful to my committee members, Professor Jeonghun Park and Professor Jemin Lee, for their insightful feedback and support.

I would like to express my appreciation to all the members of the RAMO laboratory for their support and friendship. I am especially grateful to
 혜린 for assisting me with my research and for helping me navigate the nuances of living in Korea. I would also like to thank 민규, 수진, 최진혁, 
 김진혁, and 혜성 for their help during various stages of my research. My thanks extend to all RAMO members, including 여선, 박지, 명진, 중호, 이지, 윤서, 
 Changjie, Dr. 오승은, and Dr. Arman Ahmadian, for their invaluable assistance and companionship throughout my time in the lab. Their camaraderie 
 made my experience not only productive but truly enjoyable. I will forever cherish the memories I made with the incredible people I met through RAMO.

Next, I would like to express my gratitude to the Government of the Republic of Korea and the Ministry of Education for giving me the opportunity to 
pursue my studies in Korea through the Global Korea Scholarship (GKS) program.

Finally, I would like to express my deepest gratitude to my parents, grandmother, sister, Tharindu, and friends. Though they were far away, their 
unwavering support, love, and belief in me have been a constant source of strength. I am truly thankful for their presence in my life.

\pagenumbering{gobble}
\thispagestyle{empty}
\clearpage
